Modelling state-switching functional data with hidden Markov models: supplementary material (simulation study)

### 2021-06-18 Report/Supp_Sim.tex

    	\noindent Figure \arabic{fignum}: Acceleration data for a five dives of a selected simulated data set. The raw accelerometer data is denoted as $Y^*_{t,\tilde t^*}$ which is recorded at 50 hertz while $\Zone_{t,\tilde t^*}$ and $\Ztwo_{t,\tilde t^*}$ are recorded at 0.5 hertz. Dives are separated by vertical black lines. The colour of the curve corresponds to the true fine-scale state, while the colour of the background corresponds to the true dive type. Subdive state 1 can plausibly be interpreted as ``gliding" while subdive state 2 can plausibly be interpreted as active swimming, or ``fluking".
    	\addtocounter{fignum}{1}

@@

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distribution of estimates of dive duration parameters ($\hat \mu$ and $\hat \sigma$) for each dive type. The red star represents the true values of $\mu$ and $\sigma$ for both dive types. These estimates are for the CarHHMM-DFT.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distribution of estimates of dive duration parameters ($\hat \mu$ and $\hat \sigma$) for each dive type. The red star represents the true values of $\mu$ and $\sigma$ for both dive types. These estimates are for the HHMM-DFT.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distribution of estimates of dive duration parameters ($\hat \mu$ and $\hat \sigma$) for each dive type. The red star represents the true values of $\mu$ and $\sigma$ for both dive types. These estimates are for the CarHHMM.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distribution of estimates of dive duration parameters ($\hat \mu$ and $\hat \sigma$). The red star represents the true values of $\mu$ and $\sigma$ for both dive types. These estimates are for the CarHMM-DFT. The CarHMM-DFT assumes that there is only one dive type. There is no red star which represents the true values because the CarHMM-DFT assumes that there is only one dive type.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distribution of estimates of acceleration parameters ($\hat \mu^*_A$, $\hat \sigma^*_A$, and $\hat \phi^*_A$) for each subdive state. The red star represents the true values of $\mu^*_A$, $\sigma^*_A$, and $\phi^*_A$ for both subdive states. These estimates are for the CarHHMM-DFT.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distribution of estimates of acceleration parameters ($\hat \mu^*_A$ and $\hat \sigma^*_A$) for each subdive state. The red star represents the true values of $\mu^*_A$ and $\sigma^*_A$ for both subdive states. These estimates are for the HHMM-DFT. Note that the HHMM-DFT assumes no auto-correlation in $\Zone_{t,\tilde t^*}$, so $\hat \phi_A^{*(\cdot,i^*)}$ is absent from these plots.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distribution of estimates of acceleration parameters ($\hat \mu^*_A$, $\hat \sigma^*_A$, and $\hat \phi^*_A$) for each subdive state. The red star represents the true values of $\mu^*_A$, $\sigma^*_A$, and $\phi^*_A$ for both subdive states. These estimates are for the CarHHMM. 
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distribution of estimates of acceleration parameters ($\hat \mu^*_A$, $\hat \sigma^*_A$, and $\hat \phi^*_A$) for each subdive state. The red star represents the true values of $\mu^*_A$, $\sigma^*_A$, and $\phi^*_A$ for both subdive states. These estimates are for the CarHMM-DFT.
        \addtocounter{fignum}{1}

@@

                            &                               & 2                                & $505.104$                         & $-0.686$ $(p=0.394)$          & $17.936$                             & $9.740 \pm 0.814$                             \\
                            &                               & 3                                & $9787.146$                         & $17.166$ $(p=0.718)$          & $1060.450$                             & $280.244 \pm 97.120$                             \\

@@

                            &                               & 2                                & $514.207$                         & $-2.473$ $(p=0.010)$          & $21.427$                             & $11.410 \pm 0.947$                             \\
                            &                               & 3                                & $14429.732$                         & $-32.818$ $(p=0.647)$          & $1596.020$                             & $302.698 \pm 165.935$                             \\ \hline

@@

                            &                               & 2                                & $500.928$                         & $-4.862$ $(p=0.000)$          & $24.024$                             & $10.754 \pm 1.205$                             \\
                            &                               & 3                                & $6840.995$                         & $-2928.985$ $(p=0.000)$          & $1164.288$                             & $198.162 \pm 74.720$                             \\

@@

                            &                               & 2                                & $513.263$                         & $-3.417$ $(p=0.001)$          & $22.596$                             & $12.497 \pm 1.301$                             \\
                            &                               & 3                                & $11458.795$                         & $-3003.755$ $(p=0.000)$          & $1738.793$                             & $302.698 \pm 193.611$                             \\ \hline

@@

                            &                               & 2                                & $503.762$                         & $-2.028$ $(p=0.013)$          & $18.258$                             & $9.641 \pm 0.916$                             \\
                            &                               & 3                                & $9817.785$                         & $47.805$ $(p=0.328)$          & $1089.110$                             & $280.244 \pm 136.259$                             \\

@@

                            &                               & 2                                & $515.228$                         & $-1.452$ $(p=0.125)$          & $21.087$                             & $11.371 \pm 0.990$                             \\
                            &                               & 3                                & $14437.867$                         & $-24.683$ $(p=0.740)$          & $1660.764$                             & $331.589 \pm 165.935$                             \\

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distribution of estimates of wiggliness parameters ($\hat \mu^*_W$ and $\hat \sigma^*_W$) for each subdive state. The red star represents the true values of $\mu^*_W$ and $\sigma^*_W$ for both subdive states. These estimates are for the CarHHMM-DFT.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distribution of estimates of wiggliness parameters ($\hat \mu^*_W$ and $\hat \sigma^*_W$) for each subdive state. The red star represents the true values of $\mu^*_W$ and $\sigma^*_W$ for both subdive states. These estimates are for the HHMM-DFT.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distribution of estimates of wiggliness parameters ($\hat \mu^*_W$ and $\hat \sigma^*_W$) for each subdive state. The red star represents the true values of $\mu^*_W$ and $\sigma^*_W$ for both subdive states. These estimates are for the CarHMM-DFT.
        \addtocounter{fignum}{1}

@@

        \newpage
        \subsection{CarHHMM-DFT}
        \begin{center}
        \includegraphics[width=2.5in]{../Plots/hhmm_FV_Gamma_density_-1_row_-1.png}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distributions $\hat \Gamma$, $\hat \Gamma^{*(1)}$, and $\hat \Gamma^{*(2)}$ for the CarHHMM-DFT. There are two fine-scale probability transition matrices- one corresponding to each dive type. We only plot the off-diagonal elements of the probability transition matrices since this uniquely defines a stochastic matrix. The red star represents the true values and the white star represents the mean of all maximum likelihood estimates.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distributions $\hat \Gamma$, $\hat \Gamma^{*(1)}$, and $\hat \Gamma^{*(2)}$ for the HHMM-DFT. There are two fine-scale probability transition matrices- one corresponding to each dive type. We only plot the off-diagonal elements of the probability transition matrices since this uniquely defines a stochastic matrix. The red star represents the true values and the white star represents the mean of all maximum likelihood estimates.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distributions $\hat \Gamma$, $\hat \Gamma^{*(1)}$, and $\hat \Gamma^{*(2)}$ for the CarHHMM. There are two fine-scale probability transition matrices- one corresponding to each dive type. We only plot the off-diagonal elements of the probability transition matrices since this uniquely defines a stochastic matrix. The red star represents the true values and the white star represents the mean of all maximum likelihood estimates.
        \addtocounter{fignum}{1}

>>>>

    	\noindent Figure \arabic{fignum}: Acceleration data for a five dives of a selected simulated data set. The raw accelerometer data is denoted as $Y^*_{t,\tilde t^*}$ and recorded at 50 times per second while $\Zone_{t,\tilde t^*}$ and $\Ztwo_{t,\tilde t^*}$ are recorded at once every two seconds. Dives are separated by vertical black lines. The colour of the curve corresponds to the true fine-scale state, while the colour of the background corresponds to the true dive type. Subdive state 1 can plausibly be interpreted as ``gliding", subdive state 2 can plausibly be interpreted as active swimming, and subdive state 3 can be interpretted as sudden or aggressive movement.
    	\addtocounter{fignum}{1}

@@

        \newpage

@@

        \noindent Figure \arabic{fignum}: Kernel density estimates of the distributions of estimates of dive duration parameter estimates ($\hat \mu_Y$ and $\hat \sigma_Y$). These estimates are associated with the CarHHMM-DFT and correspond to both dive types. The red star represents the true values of $\mu_Y$ and $\sigma_Y$ for both dive types.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distributions of estimates of dive duration parameter estimates ($\hat \mu_Y$ and $\hat \sigma_Y$). These estimates are associated with the HHMM-DFT and correspond to both dive types. The red star represents the true values of $\mu_Y$ and $\sigma_Y$ for both dive types.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distributions of dive duration parameter estimates ($\hat \mu_Y$ and $\hat \sigma_Y$). These estimates are associated with the CarHHMM and correspond to both dive types. The red star represents the true values of $\mu_Y$ and $\sigma_Y$ for both dive types.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distributions of dive duration parameter estimates ($\hat \mu_Y$ and $\hat \sigma_Y$). These estimates are associated with the CarHMM-DFT. There is no red star which represents the true values because the CarHMM-DFT assumes that there is only one dive type.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distributions of estimates of acceleration parameters ($\hat \mu^*_A$, $\hat \sigma^*_A$, and $\hat \phi^*_A$) for each subdive state. The red star represents the true values of $\mu^*_A$, $\sigma^*_A$, and $\phi^*_A$ for both subdive states. These estimates are for the CarHHMM-DFT.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distributions of estimates of acceleration parameters ($\hat \mu^*_A$ and $\hat \sigma^*_A$) for each subdive state. The red star represents the true values of $\mu^*_A$ and $\sigma^*_A$ for both subdive states. These estimates are for the HHMM-DFT. Note that the HHMM-DFT assumes no auto-correlation in $\Zone_{t,\tilde t^*}$, so $\hat \phi_A^{*(\cdot,i^*)}$ is absent from these plots.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distributions of estimates of acceleration parameters ($\hat \mu^*_A$, $\hat \sigma^*_A$, and $\hat \phi^*_A$) for each subdive state. The red star represents the true values of $\mu^*_A$, $\sigma^*_A$, and $\phi^*_A$ for both subdive states. These estimates are for the CarHHMM. 
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distributions of estimates of acceleration parameters ($\hat \mu^*_A$, $\hat \sigma^*_A$, and $\hat \phi^*_A$) for each subdive state. The red star represents the true values of $\mu^*_A$, $\sigma^*_A$, and $\phi^*_A$ for both subdive states. These estimates are for the CarHMM-DFT.
        \addtocounter{fignum}{1}

@@

                            &                               & 2                                & $505.10$                         & $-0.69$ $(p=0.394)$          & $17.94$                             & $9.74 \pm 0.81$                             \\
                            &                               & 3                                & $9787.1$                         & $17.2$ $(p=0.718)$           & $1060.5$                            & $280.2 \pm 97.1$                             \\

@@

                            &                               & 2                                & $514.21$                         & $-2.47$ $(p=0.010)$          & $21.43$                             & $11.41 \pm 0.95$                             \\
                            &                               & 3                                & $14430$                          & $-33$ $(p=0.647)$            & $1596$                              & $303 \pm 166$                             \\ \hline

@@

                            &                               & 2                                & $500.93$                         & $-4.86$ $(p=0.000)$          & $24.02$                             & $10.75 \pm 1.21$                             \\
                            &                               & 3                                & $6841.0$                         & $-2929.0$ $(p=0.000)$        & $1164.3$                            & $198.2 \pm 74.7$                             \\

@@

                            &                               & 2                                & $513.26$                         & $-3.42$ $(p=0.001)$          & $22.60$                             & $12.50 \pm 1.30$                             \\
                            &                               & 3                                & $11459$                          & $-3004$ $(p=0.000)$          & $1739$                              & $303 \pm 194$                             \\ \hline

@@

                            &                               & 2                                & $503.76$                         & $-2.03$ $(p=0.013)$          & $18.26$                             & $9.64 \pm 0.92$                             \\
                            &                               & 3                                & $9817.8$                         & $47.8$ $(p=0.328)$          & $1089.1$                             & $280.2 \pm 136.3$                             \\

@@

                            &                               & 2                                & $515.23$                         & $-1.45$ $(p=0.125)$          & $21.09$                             & $11.37 \pm 0.99$                             \\
                            &                               & 3                                & $14437$                          & $-25$ $(p=0.740)$          & $1661$                             & $332 \pm 166$                             \\

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distributions of estimates of wiggliness parameters ($\hat \mu^*_W$ and $\hat \sigma^*_W$) for each subdive state. The red star represents the true values of $\mu^*_W$ and $\sigma^*_W$ for both subdive states. These estimates are for the CarHHMM-DFT.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distributions of estimates of wiggliness parameters ($\hat \mu^*_W$ and $\hat \sigma^*_W$) for each subdive state. The red star represents the true values of $\mu^*_W$ and $\sigma^*_W$ for both subdive states. These estimates are for the HHMM-DFT.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distributions of estimates of wiggliness parameters ($\hat \mu^*_W$ and $\hat \sigma^*_W$) for each subdive state. The red star represents the true values of $\mu^*_W$ and $\sigma^*_W$ for both subdive states. These estimates are for the CarHMM-DFT.
        \addtocounter{fignum}{1}

@@

        \newpage

        \subsection{CarHHMM-DFT}
        \begin{center}
        \includegraphics[width=2.5in]{../Plots/hhmm_FV_Gamma_density_-1_row_-1.png}

@@

        \noindent Figure \arabic{fignum}: Kernel density  estimates of the distributions of $\hat \Gamma$, $\hat \Gamma^{*(1)}$, and $\hat \Gamma^{*(2)}$ for the CarHHMM-DFT. There are two fine-scale probability transition matrices- one corresponding to each dive type. We only plot the off-diagonal elements of the probability transition matrices since this uniquely defines a stochastic matrix. The red star represents the true values and the white star represents the mean of all maximum likelihood estimates.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density  estimates of the distributions of $\hat \Gamma$, $\hat \Gamma^{*(1)}$, and $\hat \Gamma^{*(2)}$ for the HHMM-DFT. There are two fine-scale probability transition matrices- one corresponding to each dive type. We only plot the off-diagonal elements of the probability transition matrices since this uniquely defines a stochastic matrix. The red star represents the true values and the white star represents the mean of all maximum likelihood estimates.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density  estimates of the distributions of $\hat \Gamma$, $\hat \Gamma^{*(1)}$, and $\hat \Gamma^{*(2)}$ for the CarHHMM. There are two fine-scale probability transition matrices- one corresponding to each dive type. We only plot the off-diagonal elements of the probability transition matrices since this uniquely defines a stochastic matrix. The red star represents the true values and the white star represents the mean of all maximum likelihood estimates.
        \addtocounter{fignum}{1}
